The corona of a fully convective star with a near-polar flare

### 2023-09-15 src/tex/ms.tex
[showyourwork] overleaf sync

\newcommand{\Tcool}{\input{output/T1.tex}}
\newcommand{\Thot}{\input{output/T2.tex}}
\newcommand{\eom}{\input{output/om_flare.tex}}
\newcommand{\eepic}{\input{output/epic_flare.tex}}
\newcommand{\Tqmean}{\input{output/Tqmean.tex}}
\newcommand{\Tfmean}{\input{output/Tfmean.tex}}
\newcommand{\FX}{\input{output/epic_flux.tex}}
\newcommand{\LX}{\input{output/epic_Lx.tex}}
\newcommand{\Lbol}{\input{output/Lbol.tex}}
\newcommand{\LXLbol}{\input{output/lxlbol}}
\newcommand{\ffdalpha}{\input{output/tess_ffd_alpha}}
\newcommand{\ffdbeta}{\input{output/tess_ffd_beta}}
\newcommand{\ffdr}{\input{output/R315}}

@@

We fitted the soft X-ray light curve using constant flux with a flare, applying the~\citet{davenport2014kepler} optical flare model that captures the fast rise and exponential decay of the flare seen with XMM-Newton. From the fit, we extracted the baseline flux against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:

@@

We then converted $ED$ to flare energy in the $0.2-2\,$keV band by multiplying with $L_X$.

For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}. We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}. From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $T_{\rm flare}=10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:

@@

@@

From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}). The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively. Assuming these temperatures as ground truth, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately. The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall. However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum. From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X$\FX, and a luminosity $L_X=$\LX in the $0.2$-$2\,$keV band. %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona. In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).

@@

For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-5\,$keV band. The energy of the corresponding OM flare is \eom\unskip. We use the 36ks observing baseline of XMM-Newton to calculate a flare rate in OM. Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of TESS' flare frequency distributions to lower energies. About one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277. However, this assumes that all optical flares above this energy in the optical data correspond to detected X-ray flares in the EPIC instruments, which is not always the case~\citep{paudel2021simultaneous}. \citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled flare energies are two orders of magnitude above the flare in this work, both in X-ray and optical. \citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result. We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely eclipsed by the star~\citep{johnstone2012soft}.

>>>>

\newcommand{\Tcool}{\input{output/T1.tex}}
\newcommand{\Thot}{\input{output/T2.tex}}
\newcommand{\eom}{\input{output/om_flare.tex}}
\newcommand{\eepic}{\input{output/epic_flare.tex}}
\newcommand{\Tqmean}{\input{output/Tqmean.tex}}
\newcommand{\Tfmean}{\input{output/Tfmean.tex}}
\newcommand{\FX}{\input{output/epic_flux.tex}}
\newcommand{\LXquiet}{\input{output/epic_Lx_quiet.tex}}
\newcommand{\LXflaring}{\input{output/epic_Lx_flaring.tex}}
\newcommand{\Lbol}{\input{output/Lbol.tex}}
\newcommand{\LXLbol}{\input{output/lxlbol}}
\newcommand{\ffdalpha}{\input{output/tess_ffd_alpha}}
\newcommand{\ffdbeta}{\input{output/tess_ffd_beta}}
\newcommand{\ffdr}{\input{output/R315}}

@@

For the OM flare, we median flux of the lightcurve as the quiescent flux $F_0$, against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:

@@

We then converted $ED$ to flare energy using the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}. We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}. From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $T_{\rm flare}=10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:

@@

For the X-ray flare, we use the flare duration $\Delta t$ marked in the bottom panel in Fig.~\ref{fig:lightcurves}, and multiply it by the flare luminosity in the EPIC data. The flare luminosity is equal to the difference between the quiescent and flaring X-ray luminosities in the 0.2-2 keV band, i.e.:

\begin{equation}
    E_{X, \mathrm{flare}} = \left(L_{X,\mathrm{flaring}} -  L_{X,\mathrm{quiescent}}\right) \times \Delta t
    \label{eq:xrayflare}
\end{equation}

@@

From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}). The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively. Assuming these temperatures as ground truth, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately. The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall. However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum. From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2\,$keV band. %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona. In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).

@@

For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2\,$keV band, using $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip in Eq.~\ref{eq:xrayflare}. The energy of the corresponding OM flare is \eom\unskip. We use the 36ks observing baseline of XMM-Newton to calculate a flare rate in OM. Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of TESS' flare frequency distributions to lower energies. About one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277. However, this assumes that all optical flares above this energy in the optical data correspond to detected X-ray flares in the EPIC instruments, which is not always the case~\citep{paudel2021simultaneous}. \citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled flare energies are two orders of magnitude above the flare in this work, both in X-ray and optical. \citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result. We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely eclipsed by the star~\citep{johnstone2012soft}.
